---
title: "Week 5 - Tidyr"
author: "Austin Chan"
date: "February 24, 2019"
output: html_document
---

#Tidying and Reformatting

###Call necessary packages

I will be working with tidyr and dplyr to do most of my data tidying and transformation as per the assignment instructions.

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

###Load data

I wrote the data exactly as it was presented in the assignment instructions and loaded it into R using `read.csv()`. The data is written in a wide format as suggested. I set `stringsAsFactors` to False because it makes it easier to wrangle the data when it isn't wrapped up into factors.

```{r}
ArrivalsDelays = read.csv("ArrivalsDelays.csv",stringsAsFactors = F)
ArrivalsDelays
```

###Remove na values and rename columns

The data has a blank row with na values across all columns, so I removed that blank row using `na.omit()`. Also, column names were missing for the airline companies and whether a flight was delayed or on time, so I used `rename()` to rename the columns.

```{r}
ArrivalsDelays = na.omit(ArrivalsDelays)
ArrivalsDelays = rename(ArrivalsDelays, Airline = X, IsOnTime = X.1)

ArrivalsDelays
```

###Fill in missing airlines

The raw table omits the airline name for the delayed flights to improve visual clarity when reading the table in a book. However, since I am doing analysis on the table, I need to reformat the `Airline` column to include the airline name for delayed flights. The code below subsets all unique non-blank airline names and fills in the Airline vector. The vector is then reattached to the dataframe.

```{r}
Airlines = ArrivalsDelays$Airline[ArrivalsDelays$Airline != ""]
AirlineVector = c(rep(Airlines[1],2),rep(Airlines[2],2))

ArrivalsDelays$Airline = AirlineVector

ArrivalsDelays
```

###Gather the data

After filling in the blank values, removing empty rows, and renaming columns, I can finally gather the data together into a tidy format. I used the `gather()` function to reorganize the data into a tidy format.

```{r}
ArrivalsDelays = gather(ArrivalsDelays,"City","Count",3:7)
ArrivalsDelays
```

#Analysis

##Which airline is more reliable to fly to specific cities?

I am interested in seeing which airline is more reliable to fly to each city. The way I am going to determine this is by calculating the rate of "on time" flights for each airline in each city. I will then compare them in a nice chart at the end.

###Calculating the percentage of successful flights

The first step in creating this new metric is to group the data by Airline and by City and then summing the "on time" and "delayed" flights to create the `TotalFlights` metric.

The next step is to join the new dataframe onto the original one. This allows the calculation between count and TotalFlights to be easy using `mutate()`.

The final step is to `mutate()` Count and TotalFlights to calculate the percentage of "on time" and "delayed" flights.

```{r}
TotalFlights = 
  ArrivalsDelays %>%
  group_by(Airline,City) %>%
  summarize(TotalFlights = sum(Count))

ArrivalsDelays = left_join(ArrivalsDelays,TotalFlights, by = c("Airline","City"))

ArrivalsDelays = mutate(ArrivalsDelays, Percentage = Count/TotalFlights)

ArrivalsDelays
```

###Making the plot

Below, I have written a little ggplot barchart comparing "on time" flight percentage in each city by airline. According to the plot, Alaska is more reliable across the board for every city.

```{r}
ggplot(filter(ArrivalsDelays, IsOnTime == "on time")) +
  geom_bar(aes(x = City, y = Percentage, fill = Airline),stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "On Time Flight Percentage by City")
```















